I can see clearly now: reinterpreting statistical significance

### 2018-06-20 doc.tex
In the middle of reworking (and anonymizing) the table

Instead, we suggest that using the language of ``statistical clarity'' would help researchers escape the trap of accepting 
the null hypothesis, reduce multiple types of misinterpretation associated with accepting the null hypothesis, and allow for
the continued use of NHST as a method of inference (e.g.
see \citealt{Abelson1997} for arguments for NHST).
The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the P value is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant.
``Clarity,'' on the other hand, is a particularly apt term for what hypothesis tests evaluate: a strong argument in favor or NHST is that testing the null hypothesis (even though we often know that it's false, and can never prove that it's true) is a proxy for asking whether we can see a difference clearly. In many cases, this comes down simply to whether we can be confident of the \emph{sign} of a difference or a correlation coefficient \cite{robinson2001past}.

Using the language of clarity makes sloppy thinking and misleading communication more difficult. 
For example, we often see sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant).
In fact, the sentence is erroneous even before we get to the implication: X et al.'s significance test provides information about their study, not about the underlying processes \cite{CanWeCiteThis}.
A better formulation would be ``X et al. did not find a significant effect of Y on Z.''
It seems likely, however, that this mistake would be harder to make using the language of clarity:  ``X et al. showed that the effect of Y on Z is unclear'' seems awkward.
Similarly, ``We did not find a clear difference in response between the control and sham groups'' seems both more colloquially accurate and harder to transform into something misleading than ``We did not find a significant difference\ldots''.
\citet{Bernardietal.2017} argued that ``...sociological and social significance are sacrificed on the altar of statistical significance''.
Describing statistical tests in terms of clarity could allow ``significant'' to reclaim its common English definition, and reduce conflation between statistical results and substantive significance.

@@

\begin{tabular}{p{7.5cm}p{7.5cm}}
     Language in published journal articles & Rewritten using the language of ``clarity'' \\
     	\hline
        \\
     \multicolumn{2}{c}{$\mathbf{p > 0.05 \neq}$ \textbf{zero effect}} \\
        Cadmium bioaccumulates after acute exposure but has no effect on locomotion or shelter-seeking behaviour in the
        invasive green shore crab & Cadmium bioaccumulates after acute exposure but its effect on locomotion or shelter-seeking 
        behaviour in the invasive green shore crab was statistically unclear \\ 
        \\
        No effect of elevated carbon dioxide on reproductive behaviors in the three-spined stickleback & The effect of elevated 
        carbon dioxide on reproductive behaviors in the three-spined stickleback was statistically unclear \\
        \\
        The finding that bee species richness showed no significant relationship with the area of semi-natural habitat is 
        surprising because bee species richness is usually more strongly influenced by landscape context than pollination.
& 
        Bee species richness is often found to be more strongly influenced by landscape context than pollination; however, 
        in the present study the relationship between bee species richness and area of semi-natural habitat was statistically 
        unclear \\ 
        \\
     \multicolumn{2}{c}{\textbf{P~values $\mathbf{> 0.05}$ should not be compared as evidence for anything}} \\
        ...yielded nonsignificant differences on measures of those skills with P values of at least 0.2, and most in the
        $0.7 \leq P \leq 0.9$ range.
& ...the differences between skills developed in treatment and control individuals
        were not statistically clear ($p > 0.05$) \\
        \\
     \multicolumn{2}{c}{\textbf{The difference between ``clear'' and ``not clear'' is ``not clear''}} \\
        Substrate choice correlated significantly with food choice in males $(r_{s} = 0.32, N = 44, P < 0.05)$ but not in
        females $(r_{s} = 0.32, N = 24, NS)$. ... The authors discuss the biological reasons why there is a relationship 
        between these variables in males, but not in females & 
        Substrate choice was positively correlated with food choice in males $(r_{s} = 0.32, N = 44, P < 0.05)$, but this
        relationship was not statistically clear in females $(r_{s} = 0.32, N = 24, NS)$. A biological description comparing 
        the correlation in males vs females is inappropriate without a direct test (in this case with the exact same 
        correlation estimate, this is not a sensible test).
\\
        \\
        ...risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
        (adjusted OR 2.99, 95\% CI 1.33–6.74), but no association was found between HCV coinfection and low BMD in men 
        (adjusted OR 1.26, 95\% CI 0.75–2.10). ... While the authors do not directly compare men and women they do state: 
        The precise mechanisms for the association between viral hepatitis and low BMD in HIV-infected women but not men 
        remain unclear.
& 
        ..risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
        (adjusted OR 2.99, 95\% CI 1.33–6.74), but the association between HCV coinfection and low BMD in men was not
        statistically clear (adjusted OR 1.26, 95\% CI 0.75–2.10). ... Pursuing the biological importance of HIV on BMD in 
        women and not men is premature given these statistics and requires further study.
        \\
\end{tabular}
   \captionof{table}{Rewriting misinformed language in the published literature using our proposed updated language.}

>>>>

Instead, we suggest that using the language of ``statistical clarity'' would help researchers escape the trap of accepting 
the null hypothesis, reduce multiple types of misinterpretation associated with accepting the null hypothesis, and allow for
the continued use of NHST as a method of inference (e.g.
see \citealt{Abelson1997} for arguments for NHST).
The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the P value is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant.
``Clarity,'' on the other hand, is an apt term for what hypothesis tests evaluate: a strong argument in favor of NHST is that testing the null hypothesis (even though we often know that it's false, and can never prove that it's true) is a proxy for asking whether we can see a difference clearly. In many cases, this comes down simply to whether we can be confident of the \emph{sign} of a difference or a correlation coefficient \cite{robinson2001past}.

Using the language of clarity makes sloppy thinking and misleading communication more difficult. 
For example, we often see sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant).
In fact, the sentence is erroneous even before we get to the implication: X et al.'s significance test provides information about their study, not about the underlying processes \cite{CanWeCiteThis}.
A better formulation would be ``X et al. did not find a significant effect of Y on Z.''
This mistake seems harder to make using the language of clarity:  ``X et al. showed that the effect of Y on Z is unclear'' is awkward.
Similarly, ``We did not find a clear difference in response between the control and sham groups'' seems both more colloquially accurate and harder to transform into something misleading than ``We did not find a significant difference\ldots''.
\citet{Bernardietal.2017} argued that ``...sociological and social significance are sacrificed on the altar of statistical significance''.
Describing statistical tests in terms of clarity could allow ``significant'' to reclaim its common English definition, and reduce conflation between statistical results and substantive significance.

@@

\setlength\tabcolsep{1cm}.
\begin{tabular}{p{7.0cm}p{7.0cm}}
Language from published articles & Rewritten using ``clarity'' \\
\hline\\

\multicolumn{2}{c}{$\mathbf{p > 0.05 \neq}$ \textbf{zero effect}} \\

Toxins accumulate after acute exposure but have no effect on behaviour
& Toxins accumulate after acute exposure but effects on behaviour are statistically unclear
\\

There was no effect of elevated carbon dioxide on reproductive behaviors 
& The effect of elevated carbon dioxide on reproductive behaviors was statistically unclear
\\

The finding that species richness showed no significant relationship with the area of available habitat is surprising because richness is usually strongly influenced by landscape context 
& Although species richness is usually strongly influenced by landscape context, in the present study this relationship was not statistically clear 
\\ \\

\multicolumn{2}{c}{\textbf{Comparing P~values}}
\\

\ldots\ differences between treatment and control groups were nonsignificant, with P values of at least 0.3, and most in the range $0.7 \leq P \leq 0.9$.
& \ldots\ differences between treatment and control groups were not statistically clear ($P > 0.05$)
\\ \\

\multicolumn{2}{c}{\textbf{The difference between ``clear'' and ``not clear'' is ``not clear''}}
\\

This correlation was significant in males ($\rho=0.35$, P \textless 0.05) but not females ($\rho=0.35$, NS). \ldots [Afterwards the assumption is that a difference between males and females has been demonstrated]
& Although males and females show the same correlation coefficient ($\rho=0.35$), the sign of the coefficient is statistically clear only in males  \ldots [This phrasing may suggest to the authors that confidence intervals are called for.] 
\\

...risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but no association was found between HCV coinfection and low BMD in men 
(adjusted OR 1.26, 95\% CI 0.75–2.10). ... While the authors do not directly compare men and women they do state: 
The precise mechanisms for the association between viral hepatitis and low BMD in HIV-infected women but not men 
remain unclear.
& 
..risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but the association between HCV coinfection and low BMD in men was not
statistically clear (adjusted OR 1.26, 95\% CI 0.75–2.10). ... Pursuing the biological importance of HIV on BMD in 
women and not men is premature given these statistics and requires further study.
\\
\end{tabular}
   \captionof{table}{Rewriting misinformed language in the published literature using our proposed updated language.}
